### 2015-03-24 contents/evaluation.tex
Algorithms

To evaluate the quality of model,

\subsection{Using Sparse Point Annotation}

\subsection{Using Drawing Annotation}

\clearpage