### 2015-04-10 contents/evaluation.tex
Update Abstract & Evalutation Structure

To evaluate the quality of model, 
\subsection{Models Qualiative Analysis}

\subsubsection{Model Compactness}

\subsubsection{Shape Reconstruction}

\subsection{Evaluation on Fitting}

\subsubsection{Using Sparse Point Annotation}

\subsubsection{Using Drawing Annotation}

\clearpage