### 2015-04-10 contents/evaluation.tex
added figures for evaluation section

We perform evaluation of dense deformable model
\subsection{Models Qualitative Analysis}

\subsubsection{Model Compactness}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/aam_ear_eigen_value}
        \caption{AAM Principle Components Contribution}
        \label{}
    \end{subfigure}
    ~~
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/of_ear_eigen_value}
        \caption{Shape Flow Principle Components Contribution}
    \end{subfigure}
    \caption{Principle Components Contribution}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.1\textwidth]{resources/landmark_drifting}
    \caption{Landmark Drifting}
\end{figure}

\subsubsection{Shape Reconstruction}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/warp_reconstruction_error}
        \caption{Warp Reconstruction Errors}
        \label{}
    \end{subfigure}
    ~~
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/sparse_reconstruction_error}
        \caption{Spare Reconstruction Errors}
    \end{subfigure}
    \caption{Reconstruction Errors}
\end{figure}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/warp_projecting_error_aam}
        \caption{AAM Warp Projecting Errors}
        \label{}
    \end{subfigure}
    ~~
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/warp_projecting_error_of}
        \caption{Shape Flow Warp Projecting Errors}
    \end{subfigure}
    \caption{Projecting Errors}
\end{figure}

\subsection{Evaluation on Fitting}

\subsubsection{Using Sparse Point Annotation}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.4\textwidth]{resources/face_digo_mc}
    \caption{CED Fitting Results on Face}
\end{figure}

\subsubsection{Using Drawing Annotation}
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.4\textwidth]{resources/ear_digo_mc}
    \caption{CED Fitting Results on Ear}
\end{figure}

\clearpage