### 2015-04-23 contents/conclusion.tex
Final Review Submition

We proposed the first, to the best of our knowledge, methodology that can build a dense deformable model "in-the-wild". This methodology does not require a consistent set of landmarks, avoiding the tedious task of careful manual annotations. We have exploited state-of-the-art multiframe optical flow methodologies to establish accurate and robust dense correspondences among shapes, when represented in a landmarks-free fashion. The experimental evaluations and comparisons revealed that dense AAMs that are built on top of our methodology achieve very good performance and can work effectively even in cases of objects that cannot have consistent landmarks.