### 2015-04-28 supplementary.tex
Updates from ShareLaTeX

\input{contents/header}

% \iccvfinalcopy % *** Uncomment this line for the final submission

\input{iccv_setting}

\begin{document}

%%%%%%%%% TITLE
\title{Supplementary Materials: Constructing Statistical Deformable Models with Shape Flow}

\author{Yuxiang Zhou, Joan Alabort-i-Medina, Anastasios Roussos, Stefanos Zafeiriou\\
Imperial College London\\
180 Queen’s Gate, SW7 2AZ, London, U.K.\\
{\tt\small \{yuxiang.zhou10, ja310, troussos, s.zafeiriou\}@imperial.ac.uk}}
\maketitle
\thispagestyle{empty}

The supplementary materials provided detailed evaluation of dense Active Appearance Models (dAAMs) built with proposed pipeline of constructing dense AAMs without consistent set of landmarks. There are four experiments reported to support the quality of dAAMs built

%%%%%%%%% Appendix
\appendix
\section{Fitting Results}
\label{sec:fittingresults}

\section{Model Analysis}
\label{sec:modelanalysis}

\section{Appearance \& Shape Reconstruction}
\label{sec:reconstruct}

\section{Segmentation using Dense AAM}
\label{sec:segmentation}

{\small
\bibliographystyle{ieee}
\bibliography{bib}
}

\end{document}