### 2015-07-17 contents/conclusion.tex
Changes for Rebuttal Highlighted

\textcolor{red}{In this paper, we propose the first, to the best of our knowledge, methodology for building dense deformable models "in-the-wild". Our approach does not require objects to be annotated with consistent sets of landmarks and can make use of highly-efficient (in terms of required annotation time) curve line annotations, easing the tedious task of careful manual annotation. We have exploited state-of-the-art multiframe optical flow techniques to establish accurate and robust dense correspondences among shapes, when represented in a landmarks-free fashion via the use of the SVS representation. The experimental evaluation and comparison revealed that dense AAMs that are built on top of our methodology achieve good performance and can work effectively even for objects for which is difficult to define a consistent set of landmarks (e.g. human ears).}